---
title: "Untitled"
output: html_document
date: "2024-04-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
- **Overview of R and the RStudio Interface**
  - History and Philosophy of R
  - Navigating RStudio: Panels, Scripts, Console, and Environment
  - RStudio Projects and Workspace Management
  - Installing and Managing Packages


### R Programming Fundamentals (Light Focus)
- **Basics of R Syntax and Functions**
  - Variables and Data Types
  - Control Structures: Loops and Conditional Statements
  - Writing and Using Functions
  - Debugging and Error Handling

- **Efficient Programming Practices**
  - Code Organization and Readability
  - Introduction to R's Vectorized Operations
  - Best Practices for Speed and Performance
  - Using R Profiler for Code Optimization

### Data Structures and Data Handling in R (Light Focus)
- **Overview of Vectors, Matrices, Lists, and Data Frames**
  - Creating and Manipulating Vectors
  - Operations with Matrices and Arrays
  - Understanding Lists and Their Uses
  - Data Frames for Tabular Data

- **Data Handling Techniques with dplyr**
  - Selecting, Filtering, and Arranging Data
  - Mutating and Summarizing Data Sets
  - Joins and Data Merging Techniques
  - Group Operations with `group_by` and `summarise`

### Data Visualization in R (Light Focus)
- **Introduction to ggplot2 and Basic Plotting**
  - ggplot2 Syntax and Layering System
  - Creating Histograms, Bar Plots, and Scatter Plots
  - Aesthetics and Themes

- **Customizing Graphs for Data Presentation**
  - Customizing Axes, Legends, and Labels
  - Fine-tuning Scales and Coordinates
  - Adding Annotations and Custom Geometries

### Working with Different Types of Data (Light Focus)
- **Overview of Qualitative and Quantitative Data**
  - Defining Qualitative vs Quantitative Data
  - Sampling Strategies and Data Collection Methods
  - Structuring Data Sets for Analysis

- **Best Practices in Data Collection and Cleaning**
  - Identifying and Handling Missing Data
  - Data Type Conversions and Formatting
  - Detecting and Dealing with Outliers

### Descriptive Statistics
- **Measures of Central Tendency, Variability, and Standard Scores**
  - Calculating Mean, Median, Mode
  - Variance, Standard Deviation, and Range
  - Normalization and Standardization of Scores

- **Visualizing Distributions and Relationships**
  - Box Plots, Violin Plots, and Density Plots
  - Pairwise Relationships with Scatter Plots and Correlograms
  - Introduction to Multidimensional Scaling and PCA

### Probability Distributions and Statistical Concepts
- **Understanding and Applying Common Probability Distributions**
  - Binomial, Poisson, and Normal Distributions
  - Continuous vs Discrete Distributions
  - The Central Limit Theorem and Its Importance

- **Exploring Random Variables and Expected Values**
  - Defining and Calculating Expected Values
  - Variance and Standard Deviation of Random Variables
  - The Law of Large Numbers

### Inferential Statistics and Hypothesis Testing
- **Theoretical Basis for Hypothesis Testing and Significance**
  - Null and Alternative Hypotheses
  - Type I and Type II Errors
  - p-Values and Confidence Intervals

- **Chi-Square, t-tests, z-tests, and Non-Parametric Tests**
  - Conducting and Interpreting Chi-Square Tests
  - One-sample and Two-sample t-tests
  - Non-Parametric Alternatives to Parametric Tests

### Analysis of Variance (ANOVA)
- **Conducting and Interpreting One-Way and Two-Way ANOVA**
  - Assumptions of ANOVA
  - Interpreting ANOVA Tables
  - Post Hoc Tests and Multiple Comparisons

- **Understanding the F-test and its Applications**
  - The F Distribution and the F Ratio
  - Between-groups and Within-groups Variance
  - Applications in Experimental Design

### Regression Analysis
- **Simple and Multiple Linear Regression (OLS, MLR)**
  - Constructing and Fitting Regression Models
  - Understanding Coefficients and Predictions
  - Multiple Regression and Adjusting for Confounders

- **Diagnostics and Assumptions of Linear Models**
  - Residual Analysis and Model Fit
  - Checking for Heteroscedasticity and Multicollinearity
  - Model Selection Criteria (

AIC, BIC, R-squared)

### Categorical Data Analysis
- **Introduction to Logistic Regression**
  - Odds Ratios and Logit Function
  - Model Fitting and Interpretation
  - Assessing Model Goodness of Fit

- **Modeling and Interpretation of Binary Outcomes**
  - Predictive Modeling with Binary Data
  - Evaluating Model Performance (ROC Curves, AUC)
  - Multiclass Logistic Regression

### Advanced R Programming Techniques
- **Writing Functions and Loops for Automating Tasks**
  - Function Arguments and Return Values
  - Looping Constructs: `for`, `while`, and `apply` Family
  - Avoiding Loops: Vectorization and Parallel Processing

- **Advanced Data Manipulation and Transformation**
  - Tidyverse Approaches to Data Wrangling
  - Dealing with Text and Categorical Data
  - Working with Dates and Times

### Data Import and Export
- **Techniques for Efficient Data Import and Export**
  - Reading and Writing CSV, Excel, and JSON
  - Connecting to Databases with R
  - Using APIs for Data Retrieval

- **Working with Different File Formats and Databases**
  - Understanding File Formats and Their Uses
  - Data Storage and Retrieval Best Practices
  - Introduction to SQL for Data Analysis

### Bivariate and Multivariate Analysis
- **Exploratory Data Analysis Techniques**
  - Summary Statistics and Initial Data Checks
  - Exploring Relationships with Crosstabs and Mosaic Plots
  - Cluster Analysis and Discriminant Analysis

- **Correlation, Causation, and Regression Techniques**
  - Pearson and Spearman Correlation Measures
  - Path Analysis and Structural Equation Modeling
  - Multivariate Regression Models

### Practical Applications in R
- **Case Studies and Real-World Data Analysis**
  - Sector-specific Analysis (e.g., Healthcare, Finance, Marketing)
  - Interpreting Results in a Business Context
  - Addressing Ethical Considerations in Data Analysis

- **Developing a Capstone Project to Apply Learned Skills**
  - Identifying a Problem and Formulating Questions
  - Data Collection and Analysis Plan
  - Presenting Findings and Recommendations

### Reporting and Communication
- **Crafting Reports with R Markdown**
  - Basics of R Markdown for Reporting
  - Dynamic Report Generation and Parameterized Reports
  - Creating Interactive Dashboards with Shiny

- **Effective Communication of Statistical Findings**
  - Visualization Techniques for Storytelling with Data
  - Translating Statistical Outputs into Business Insights
  - Communicating Uncertainty and Model Limitations



When creating R Markdown (Rmd) documents, you can embed different types of code chunks that serve various purposes. These chunks can be tailored for demonstration purposes, to show pre-run code with results, and to provide interactive code exercises. Here’s how you can handle each usage scenario in your Rmd documents, which will also work when you convert them into a Bookdown project:

### Usage 1: Code for Demonstration (Non-Executable)
To include code for demonstration purposes where the code is displayed but not run, you can use the `eval=FALSE` chunk option. This will display the code without actually executing it when the document is knit.

```r
```{r eval=FALSE}
# This is demonstration code
summary(cars$speed)
```
```

### Usage 2: Pre-Run Code Showing Results
To display code along with its results without allowing it to be changed, simply include the code chunk without any special options. When the Rmd document is knit, the code will run, and the output will be included in the document.

```r
```{r}
# Pre-run code for results
plot(cars$speed, cars$dist)
```
```

The output, such as plots or tables, will be generated and embedded into the final document automatically.

### Usage 3: Interactive Code Writing and Execution
To allow users to write and run code within the Rmd document, you can use a standard R code chunk. Once the Rmd is knit, the code can be run in the output document if you are generating an HTML document using the `runtime: shiny` option in the YAML header. This enables interactivity in the HTML document but note that this feature won’t be available in PDFs or other non-interactive formats.

Here’s what the YAML header would look like to enable Shiny for an HTML document:

```yaml
---
title: "Interactive Book"
output: html_document
runtime: shiny
---
```

And here's how you include a code chunk that the user can interact with:

```r
```{r}
# Interactive code chunk
summary(cars$speed)
```
```

If you want to provide a terminal for students to write and execute commands, you can embed a Shiny app within your Bookdown book. However, the app must be hosted on a Shiny server or a service like shinyapps.io, and the book must be rendered as an HTML document.

### Additional Tips for Bookdown

When transitioning from Rmd files to Bookdown, keep the following in mind:

- **Chunk Naming**: It’s good practice to name your chunks uniquely to avoid duplicate label issues in Bookdown when combining multiple Rmd files.
  
- **Chunk Options**: Review the chunk options available in knitr. Options like `include=FALSE` can hide code but show its output, `echo=FALSE` will hide the code but execute it, etc.

- **Interactive Documents**: For full interactivity, like Shiny exercises at the end of topics, ensure your output format supports it (i.e., HTML).

- **External Resources**: For exercises, you might want to consider external resources like DataCamp’s Workspace or other interactive platforms where students can practice.

- **Session Management**: Manage the R session carefully for reproducibility. Bookdown compiles documents in a fresh R session by default, which is usually what you want for consistency.

- **Output Management**: Use the `_output.yml` file to manage output options for different formats (HTML, PDF, etc.) within Bookdown.

Remember that the interactivity of code chunks that allow users to modify and run code is limited to HTML documents with Shiny capabilities and won't be interactive in PDFs or ePub formats. For a fully interactive online learning experience, you may need to complement your Bookdown content with external tools or platforms that support interactive coding.